---
title: "Heart Stroke Prediction - Code & Technical Analysis"
author: "Team KAT"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: TRUE
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r basic_libraries, include=FALSE}
library(ezids)
library(ggplot2)
library(gridExtra)
library(corrplot)
library(ROSE)
library(dplyr)
library(ggcorrplot)
```

# Loading the dataset
```{r load_csv, include=TRUE, echo=TRUE}
# reading the dataset from github repository
github_file_path <- "https://raw.githubusercontent.com/kashyapnimmagadda/DATS-6101-GROUP-KAT/main/DataSet/framingham.csv"
stroke_df <- read.csv(url(github_file_path))
stroke_df
```

```{r desc, include=TRUE, echo=TRUE}
# structure of the dataset
str(stroke_df)
```
# Data Cleaning and summary of the dataset
```{r data_cleaning, include=TRUE, echo=TRUE}

stroke_df <- stroke_df[, -which(names(stroke_df) == "education")]
stroke_df <- stroke_df[, -which(names(stroke_df) == "heartRate")]

names(stroke_df)[names(stroke_df) == "TenYearCHD"] <- "stroke"
colnames(stroke_df)[1] <- "gender"

stroke_df$gender[stroke_df$gender == "0"] <- "F"
stroke_df$gender[stroke_df$gender == "1"] <- "M"

# converting the numeric variables to factor variables
stroke_df$gender = as.factor(stroke_df$gender)
stroke_df$currentSmoker = as.factor(stroke_df$currentSmoker)
stroke_df$BPMeds = as.factor(stroke_df$BPMeds)
stroke_df$prevalentStroke = as.factor(stroke_df$prevalentStroke)
stroke_df$prevalentHyp = as.factor(stroke_df$prevalentHyp)
stroke_df$diabetes = as.factor(stroke_df$diabetes)
stroke_df$stroke = as.factor(stroke_df$stroke)
stroke_df$BMI = as.numeric(stroke_df$BMI)

# To get the summary statistics of the dataset
summary(stroke_df)

```


```{r find_na_fix, include=TRUE, echo=TRUE}
# To find the NA's in the dataset
paste("The NA's in the dataset is:",sum(is.na(stroke_df)))
```

```{r na_fix, include=TRUE, echo=TRUE}
# Replacing NA values with average BMI value
stroke_df$BMI[is.na(stroke_df$BMI)] = mean(stroke_df$BMI,na.rm = TRUE)
# Replacing NA values with average glucose value
stroke_df$glucose[is.na(stroke_df$glucose)] = mean(stroke_df$glucose,na.rm = TRUE)
# Replacing NA values with average cigsPerDay value
stroke_df$cigsPerDay[is.na(stroke_df$cigsPerDay)] = mean(stroke_df$cigsPerDay,na.rm = TRUE)
# Replacing NA values with average totChol value
stroke_df$totChol[is.na(stroke_df$totChol)] = mean(stroke_df$totChol,na.rm = TRUE)
stroke_df$BPMeds<- as.integer(stroke_df$BPMeds)
# Replacing BPMeds NA values with Unknown value
stroke_df$BPMeds <- ifelse(is.na(stroke_df$BPMeds), "Unknown", stroke_df$BPMeds)
# Replacing BPMeds NA values with mean value
stroke_df$BPMeds[is.na(stroke_df$BPMeds)] = mean(stroke_df$BPMeds, na.rm = TRUE)
#stroke_df$BPMeds <- ifelse(is.na(stroke_df$BPMeds), "Unknown", stroke_df$BPMeds)
paste("The NA's in the dataset after replacing values is:",sum(is.na(stroke_df)))
```

```{r subsets, include=TRUE, echo=TRUE}
# subsetting the data for various analyses 
stroke_1 = subset(stroke_df, stroke == 1)
stroke_0 = subset(stroke_df, stroke == 0)
stroke_1_female = subset(stroke_df, stroke == 1 & gender == "F")
stroke_1_male = subset(stroke_df, stroke == 1 & gender == "M")

# creating different columns for bmi, age and average_glucose_level based on different bucketing for each variables
dat <- within(stroke_df, {   
  BMI.cat = NA # need to initialize variable
  BMI.cat[BMI < 18.5] = "underweight"
  BMI.cat[BMI >= 18.5 & BMI < 25] = "normal"
  BMI.cat[BMI >= 25 & BMI < 30] = "overweight"
  BMI.cat[BMI >= 30 & BMI < 40] = "obesity"
  BMI.cat[BMI >=40] = "severe obesity" 
  gluc.cat = NA # need to initialize variable
  gluc.cat[glucose < 60] = "Below 60"
  gluc.cat[glucose >= 60 & glucose < 90] = "60 - 90"
  gluc.cat[glucose >= 90 & glucose < 120] = "90 - 120"
  gluc.cat[glucose >= 120 & glucose < 180] = "120 - 180"
  gluc.cat[glucose >= 180 & glucose < 273] = "180 - 273"
  age.cat = NA
  age.cat[age<=20] = "Under 20"
  age.cat[age>=21 & age<= 40 ] = "20-40"
  age.cat[age>=41 & age<= 60 ] = "40-60"
  age.cat[age>=61 & age<= 80 ] = "60-80"
  age.cat[age>=80 ] = "above 80"
   } )

dat$BMI.cat<- factor(dat$BMI.cat,levels= c("underweight","normal","overweight","obesity","severe obesity" ))
dat$gluc.cat<- factor(dat$gluc.cat,levels= c("Below 60","60 - 90","90 - 120","120 - 180","180 - 273" ))
dat$age.cat<- factor(dat$age.cat,levels= c("Under 20","20-40","40-60","60-80","above 80"))
```

```{r}
head(stroke_df)
```

#Plots
```{r variable_dist, include=TRUE, echo=TRUE}
# creating density plots to see the distribution of the numerical variables

ggplot(stroke_df, aes(x=age)) +  geom_density(fill="skyblue", color="skyblue", alpha=0.5)+ theme_bw() + theme()+ ggtitle("Distribution for age")

ggplot(stroke_df, aes(x=BMI)) +  geom_density(fill="pink", color="pink", alpha=0.5)+ theme_bw() + theme()+ ggtitle("Distribution for BMI")

ggplot(stroke_df, aes(x=glucose)) +  geom_density(fill="lightgreen", color="lightgreen", alpha=0.5)+ theme_bw() + theme()+ ggtitle("Distribution for Average Glucose Level")

# Count of Males and Females in the Dataset
ggplot(stroke_df, aes(x=gender,fill=gender))+geom_bar()+ggtitle("Count of Male and Female in the Dataset")+ theme_bw() +
  theme()+ xlab("Gender") + ylab("Count of people")
```

## Distribution of the target variable 
```{r target_var, include=TRUE, echo=TRUE}
ggplot(stroke_df, aes(x = stroke)) + 
  geom_bar(aes(fill = ..count..)) + 
  scale_fill_gradient('Count', low = "#F0E442", high = '#999999') + 
  labs(x = "stroke", y = "Count of People", title = "Distribution of Target variable (stroke)")
```

# BMI vs stroke 
```{r}
ggplot(stroke_df, aes(x=BMI, fill=stroke)) + geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity')+ theme_bw() + theme()+ggtitle("BMI vs Stroke")
```

### Does it depend based on the gender?
```{r gender_job_smoking, include=TRUE, echo=TRUE}
# seeing if gender has any affect on stroke based on smoking and work type
ggplot(stroke_1_female,aes(x=age, fill= currentSmoker))+geom_bar(position = "dodge")+ggtitle("Distribution of females who had a stroke based on their smoking habits")+ theme_bw() + theme()+ xlab("Age") + ylab("Count of people ")

ggplot(stroke_1_male,aes(x=age, fill= currentSmoker))+geom_bar(position = "dodge")+ggtitle("Distribution of males who had a stroke based on their smoking habits")+ theme_bw() + theme()+ xlab("Age") + ylab("Count of people ")
```
As seen in the graph above, the key causes that caused a stroke in females were never smoked and formerly smoked, whereas in males, the reason for stroke was formerly smoked and smokes.
```{r}
#To see if these variables are dependent we did a chi square test.

chisq.test(stroke_df$stroke, stroke_df$gender)
chisq.test(stroke_df$stroke, stroke_df$currentSmoker)
chisq.test(stroke_df$stroke, stroke_df$BPMeds)
chisq.test(stroke_df$stroke, stroke_df$prevalentStroke)
chisq.test(stroke_df$stroke, stroke_df$prevalentHyp)
chisq.test(stroke_df$stroke, stroke_df$diabetes)

# t-test for totChol
t.test(totChol ~ stroke, data = stroke_df)

# t-test for sysBP
t.test(sysBP ~ stroke, data = stroke_df)

# t-test for diaBP
t.test(diaBP ~ stroke, data = stroke_df)

# t-test for BMI
t.test(BMI ~ stroke, data = stroke_df)

# t-test for glucose
t.test(glucose ~ stroke, data = stroke_df, var.equal = FALSE)

#As we can see from the results, the p values for both tests are less than 0.05, thus we can fail to accept the null hypothesis and infer that smoking status and work type are related to stroke.
```

```{r age_stroke_1, include=TRUE, echo=TRUE}
# seeing if age has effect in stroke
#FIX
ggplot(stroke_1, aes(x = age, fill=stroke))+ geom_density(alpha = 0.3)+ ggtitle("Density plot for age of people who had stroke")+ theme_bw() + theme()
```
In general, we know that as one gets older, the likelihood of contracting an illness increases. We wanted to see if this statement held true with our dataset. The above graph clearly shows that as one's age grows, the likelihood of having a stroke increases.

### Does BMI with age cause stroke?
```{r age_bmi, include=TRUE, echo=TRUE}

# how bmi and age is stroke in people 
ggplot(stroke_1, aes(x=BMI, y=age, color=stroke))+geom_point(size=3)+ ggtitle("Scatter plot for age vs bmi for people who had a stroke")

ggplot(stroke_0, aes(x=BMI, y=age, color=stroke))+geom_point(size=3)+ggtitle("Scatter plot for age vs bmi for people who didnt have a stroke")


ggplot(stroke_1, aes(x=BMI, y=age, color=stroke))+geom_boxplot(fill = "skyblue")+ ggtitle("Box plot for age vs bmi for people who had a stroke")


ggplot(stroke_0, aes(x=BMI, y=age, color=stroke))+geom_boxplot(fill = "skyblue")+ ggtitle("Box plot for age vs bmi for people who didnt have a stroke")

```
The categories are encoded as:

- Underweight - BMI < 18.5
- Normal - BMI >= 18.5 and BMI <= 25
- Overweight - BMI >= 25.0 and BMI <= 30
- Obesity - BMI >= 30.0 and BMI < 40
- "Extreme" or Severe Obesity - BMI >= 40 

```{r age_bmi.cat, include=TRUE, echo=TRUE}

# boxplot to see if bmi along with age has anything to do with stroke
ggplot(dat, aes(x=BMI.cat,y = age, fill=stroke))+geom_boxplot()+ ggtitle("Boxplot for Age vs BMI(categorical)")

```
### Does age with average glucose level affect stroke?
```{r age_avg_glucose, include=TRUE, echo=TRUE}
# seeing if average glucose level with age has an effect on stroke
stroke_1_dat = subset(dat, stroke==1 )
ggplot(stroke_1_dat, aes(fill=stroke, y = age, x = gluc.cat))+ geom_boxplot()+ ggtitle("Boxplot for Age vs Average Gluclose level(Categorical) for people with stroke")
```
### Does age cause hypertension?

```{r age_hypertension, include=TRUE, echo=TRUE}
# seeing if a person can get hypertension with age 
ggplot(data = stroke_df, aes(x=as.character(prevalentHyp), y=age, fill=stroke)) +
    geom_boxplot() +
    labs(title="Age distribution by hypertension", x="hypertension", y="age")
```

```{r}
ggplot(dat, aes(x = BMI.cat, fill = stroke)) + 
  geom_bar() + 
  labs(x = "BMI Category", y = "Count", title = "Distribution of Stroke Cases by BMI Category")

```

```{r}
ggplot(stroke_df, aes(x = age, fill = stroke)) +
  geom_histogram(binwidth = 5, position = "dodge") +
  labs(x = "Age", y = "Count", title = "Distribution of Age by Stroke Status") +
  scale_x_continuous(breaks = seq(0, 100, by = 5))

```

```{r}
ggplot(stroke_df, aes(x = BMI, y = glucose, color = stroke)) +
  geom_point(alpha = 0.7) +
  labs(x = "BMI", y = "Glucose Level", title = "Relationship between BMI and Glucose Levels") +
  scale_color_manual(values = c("gray", "red"))

```

```{r}
ggplot(dat, aes(x=sysBP, y=diaBP, color=age)) +
  geom_point() +
  labs(title="Scatter plot of systolic vs diastolic blood pressure", x="Systolic BP", y="Diastolic BP", color="Age") +
  scale_color_gradient(low="blue", high="red")

```

```{r}
ggplot(dat, aes(x = gluc.cat, fill = stroke)) + 
  geom_bar() + 
  ggtitle("Ten-Year CHD and Glucose Category Stacked Barplot")
```
